\LARGE Untangling the link between traits, size and growth rate in plants

### 2016-09-20 ms/ms.tex
Full draft
- Add abstract, conclusions, minor edits throughout
- Improve figure and table captions
- Add Remko as coauthor
- Add tables in SuppMat, details about hyper-parameterisation
- Better table of contents in SuppMat

\keywords{growth strategy, life-history,  ontogeny, size, leaf mass per area, wood density, shade-tolerance, growth rate}

@@

\section{Abstract}
We present a new theoretical framework for understanding the effect of diverse functional traits on plant growth.

By disentangling the effects of plant size, light environment and traits on plant growth, we hope to provide a solid theoretical foundation linking trait-based trade-offs to demography and thus provide a platform for understanding growth across diverse species around the world.

@@

One possible reason that theoretical predictions have not been supported is that the effects of traits in existing theory are realised mainly via influences on mass production \citep{Enquist-2007}. By contrast, the physiology of traits such as leaf mass per unit leaf area and wood density suggests they have influence by altering the allocation of energy among different tissues, rather than changing mass production \citep{Falster-2011, Gibert-2016}. A second concern for theory focussed on mass production is that altered mass production is only possible outcome. Measuring mass production is really only practical for small plants that can be easily harvested. For larger plants, growth is more-often measured as increment in either diameter or height. For example, research centres and forestry agencies have been recording stem diameter growth in 1000's of plots across the globe \citep{Purves-2008, Anderson-2015, Kunstler-2016}. Theory focussing on mass production is therefore of limited potential use.

In this paper we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth (Table \ref{tab:phenomena}); and thereby offer fresh insight into the way traits influence plant demography throughout across the life-cycle. This model can be viewed as an integration of several related approaches \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011}. Our primary focus in this article is to explain the pattern that has been gradually emerging: that the effect of traits on plant growth may be modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}. As noted above, the paper by \citet{Gibert-2016} outlined a framework describing how traits influence growth. Here we extend the results of \citet{Gibert-2016} to provide a functioning model and explain, from the point of view of energy uptake and allocation, why it makes sense that the effect of some traits changes with size. We also show how our approach can account for other phenomena, including changes in growth and shade tolerance with traits, individual size, and light environment. We submit that, trait-based approaches -- which focus on differences among species -- should seek to be integrated within a general model of plant growth, also able to capture patterns of growth through ontogeny. Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Herault-2011}. In contrast, basal-area growth continues to increase with size \citep{Sillett-2010, Stephenson-2014}. And all growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}. Shade tolerance also varies among species, correlates with traits \citep{Lusk-2008,Poorter-2006}, and tends to decrease with increasing size \citep{Givnish-1988, Lusk-2008}. These diverse phenomena -- summarised in Table \ref{tab:phenomena} -- are deserving of a comprehensive and integrated explanation.

@@

We assume growth is fundamentally driven by energy income and subsequent distribution throughout the plant. Applying a standard approach, the amount of biomass available for growth, $\frac{{\rm d}B}{{\rm d}}t$ is given by the difference between income (photosynthesis) and losses (respiration and turnover) within the plant \citep{Makela-1997, Thornley-2000} (Fig. \ref{fig:conceptual}a, eqn \ref{eq:dbdt}). Photosynthesis is the product of the average photosynthetic rate per unit leaf area ($\bar{p}$) and total leaf area ($A_{\rm l}$). The constant $y$ in eqn \ref{eq:dbdt} converts mol CO$_2$ into units of biomass. We assume that $\bar{p}$ increases with light but with diminishing returns, and respiration and turnover rates of different tissues are constants that might differ with traits (see below). While many vegetation models use a detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$, e.g. as functions of temperature, such detail is not needed here, as even with detailed models, $\frac{{\rm d}B}{{\rm d}}t$ always comes down to simple subtraction of income and losses.

@@

Eqn \ref{eq:dbdt} can also be used to estimate a measure of shade tolerance, sometimes referred to as the the ``whole-plant-light-compensation point'' ({\wplcp}): the  the light level at which a plant's photosynthetic gains just balance the costs of tissue turnover and respiration \citep{Givnish-1988, Baltzer-2007, Lusk-2013}. In general, average assimilation rate per leaf area $\bar{p}$ is expected to increase with light level or canopy openness, $E$. The {\wplcp} can be estimated by solving the value $E=E^*$ giving $\frac{{\rm d}B} {{\rm d}t}(E)=0$. From eqn \ref{eq:dbdt}, this occurs when
\begin{equation}\label{eq:wplcp}
\bar{p}(E^*) =\frac{\sum_{i={\rm l,b,s,r}} \, M_{\rm i} \left(\frac{k_{\rm i}}{y} + r_{\rm i}\right)}{A_{\rm l}},
\end{equation}
where $\bar{p}(E)$ is the rate of photosynthesis at canopy openness $E$. The {\wplcp} occurs at the points where the photosynthetic production (per unit leaf area) line intersects with the sum of maintenance and respiration costs (per unit leaf area) for each tissue (Fig. \ref{fig:wplcp_idea}). Traits will influence the {\wplcp} if they effect either income or costs. Also, since the amount of stem support naturally increases with height, the {\wplcp} also naturally increases with height \citep{Givnish-1988} (Fig. \ref{fig:wplcp_idea}).

@@

It worth noting that because eqns \ref{eq:dMtdt}-\ref{eq:dDdt} are derived derived using the chain rule of differential calculus, they are mathematically true, and must therefore hold for a wide range of potential growth models. To make explicit predictions via this framework, however, requires a explicit model of plant construction and function; i.e. we must put equations behind all there terms in Fig. \ref{fig:conceptual}b.

Here we adopt a model of plant construction and function that can be considered a first-order functional-balance model, similar to those implemented by \citet{Makela-1997} and \citet{Moorcroft-2001}. We could also call it ``isometric'', because the assumptions preserve the proportions among the relevant variables as the plant grows. Table \ref{tab:allometry} provides key equations of the model (see Supplementary Material \ref{sec:ff16} or \citet{Falster-2016} for full derivation). In particular, we assume:
\begin{description}
\item[Assumption 1:] A log-log scaling relationship with slope of 0.5 between the leaf area of a plant and its height (eqn \ref{eq:ha}).
\item[Assumptions 2:] A log-log scaling relationship with slope of 1.0 between the leaf area and cross-sectional area of sapwood in a plant (eqn \ref{eq:as}).
\item[Assumptions 3:] A log-log scaling relationship with slope of 1.0 between the leaf area and cross-sectional area of bark and phloem in a plant (eqn \ref{eq:ab}).
\item[Assumptions 4:] A log-log scaling relationship with slope of 1.0 between the leaf area and mass of fine roots in a plant (eqn \ref{eq:mr}).
\item[Assumption 5:] The vertical distribution of leaf within the canopy of an individual plant remains constant through ontogeny.
\end{description}
Assumption 1 accounts for the architectural layout of the plant. Assumptions 2-4 are realisations of the pipe model \citep{Shinozaki-1964}, whereby the cross-sectional area of conducting tissues are proportional to leaf area. To describe the vertical distribution of leaf area within the canopy of an individual plant (assumption 5), we use the model of \citet{Yokozawa-1995}, which can account for a variety of canopy profiles through a single parameter $\eta_c$, varying from 0-1 (for details, see Supplementary Materials).

Combined, the functional-balance assumptions from Table \ref{tab:allometry}a lead directly to equations describing the mass of sapwood and bark in relation to leaf, and the amount of leaf in relation to height, described in Table \ref{tab:allometry}b. Substituting from Table \ref{tab:allometry} into eqns \ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives all the necessary terms needed to implement the growth model described in Fig. \ref{fig:conceptual}.

An important prediction arising from the eqns in Table \ref{tab:allometry} is that the amount of live stem tissue supporting each unit of leaf area increases linearly with  height:
\begin{equation}\label{eq:msml}
\frac{M_{\rm b} + M_{\rm s}}{A_{\rm l}} = (1+ \alpha_{\rm b1}) \,\theta \, \rho \, \eta_c \, H,
\end{equation}
where $\alpha_{\rm b1}, \theta, \rho, \eta_c$ are all traits, i.e. properties that are assumed approximately constant through ontogeny.

@@

We now consider how trait-based trade-offs enter into the above growth model. It is essential that any trait include both a benefit and cost in terms of plant function and/or life history; otherwise we would expect never-ending directional trait values towards more beneficial values. For present purposes, we consider five prominent traits for which we can posit specific costs and benefits, outlined in Table \ref{tab:traits}. In postulating potential benefits and costs, we consider only those thought to arise as direct biophysical consequences of varying a trait. The trade-offs are then implemented as follows.
\begin{description}
\item[Seed mass, $\omega$:] There is a direct energetic trade-off between the a plant's fecundity and the size of it's seeds. Moreover, we assume larger seeds result in larger seedlings; the seed-size number tradeoff thereby translates into a demographic tradeoff between the number of seedlings produced by a parent and their initial size.

\item[Height at maturation, $H_{\rm m}$:] This trait moderates the inevitable energetic trade-off between growth and reproduction. We assume the fraction of energy allocated to reproduction, $1- \frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ is determined by $H_{\rm m}$ via a function $r(H, H_{\rm m}) = \frac{r_{\rm r1}} {1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{\rm m}\right)\right)}$, where $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition.
% By default, we set $r_{\rm r1}=0.8$ and $r_{\rm r2}=10$, implying a relatively rapid transition to reproduction at $H=H_{\rm m}$ (Fig. \ref{fig:S3}).

\item[Nitrogen per unit leaf area,  $\nu$:] We allow for the maximum photosynthetic capacity of the leaf to vary with leaf nitrogen per unit area, as $A_{\rm max} = \beta_{{\rm lf1}} \, \left(\frac{\nu}{\nu_0}\right)^{\beta_{{\rm lf5}}}$, where $\beta_{{\rm lf1}}$, $\nu_0$ and $\beta_{{\rm lf5}}$ are constants. Respiration rates per unit leaf area are also assumed to vary linearly with leaf nitrogen per unit area, as $\beta_{{\rm lf4}}\, \nu$.\footnote{Add plot of this relationship, from data in \citet{Wright-2004}.}

\item[Leaf mass per unit area, $\phi$:] This trait directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}). In addition, we link $\phi$ to the rate of leaf turnover ($k_{\rm l}$), based on a widely observed scaling relationship from \citet{Wright-2004}: $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants. The rate of leaf respiration per unit area is assumed independent of $\phi$, as such the mass-based rate is adjusted accordingly whenever $\phi$ is varied.

\item[Wood density, $\rho$:] This trait directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}). In addition, we link $\rho$ to the rate of sapwood and bark turnover, mirroring the relationship assumed for leaf turnover: $k_{\rm b} = k_{\rm s} = \beta_{{\rm ks1}} \, \left(\frac{\rho}{\rho_0}\right)^{-\beta_{{\rm ks2}}}$ where $\beta_{{\rm ks1}}$, $\rho_0$ and $\rho_{{\rm ks2}}$ are empirical constants. The rate of sapwood and bark respiration per unit stem volume is assumed to be independent of $\rho$, as such the mass-based rate is adjusted accordingly  whenever $\rho$ is varied.
 \end{description}

@@

The growth model described above has been implemented as the \texttt{FF16} physiological model within the {\plant} package \citep{Falster-2016} for \textsc{r} \citep{R-2015}. The {\plant} package also makes use of supporting packages \texttt{Rcpp} \citep{Eddelbuettel-2013} and the Boost Library for C++\citep{Schaling-2014}, via the package \texttt{BH} \citep{Eddelbuettel-2015}. To encode the trait-based trade-offs described above, we use the capacity to in {\plant} to provide a ``hyper-parameterisation'' function, which enables various parameters to covary with traits. Parameters used in the current analysis were sourced from \citet{Falster-2016}, with the exception of XXXX (see Table \ref{tab:params} for values).

@@

The functional-balance assumptions outlined in Table \ref{tab:allometry}a are well-supported by the available data, at least to first degree. Fig. \ref{fig:assumptions} shows data sourced from the \textsc{baad} \citep{Falster-2015b}, with solid lines indicating {\sma} lines fit to each species in the data. Dashed lines indicate the slope of predictions under functional-balance assumptions in Table \ref{tab:allometry}a. As expected, species differed in elevation, but less so in the slope of the fitted lines, with slopes aligning with those predicted by the functional-balance assumptions.

@@

Our growth model suggests an intrinsically size-dependent pattern of growth, which aligns with well-known empirical patterns (Table \ref{tab:phenomena}). The panels in Fig. \ref{fig:conceptual} show the expected patterns for a typical woody plant. Height growth shows a hump-shaped pattern with size, first increasing then decreasing. This pattern results from systematic changes in the four components of eqn \ref{eq:dhdt} with size (Fig. \ref{fig:conceptual}), including a strong decline in the fraction of plant that is leaf declines with increasing size (Fig. \ref{fig:conceptual}), and increasing reproductive allocation. In contrast, basal-area growth continues to increase with size, due to an increasing influence of stem turnover. Diameter growth shows a weakly hump shaped curve, tapering off slightly at larger sizes, in part because of the allometric conversion from basal area to diameter (eqn \ref{eq:dDdt}, and in part because of increased reproductive allocation in older trees (Fig \ref{fig:conceptual}). All growth measures decrease sharply with size when expressed as relative growth rates.

@@

We analysed the response of growth rate to five different traits under the assumed trade-offs (Table \ref{tab:traits}). We considered changes in absolute and relative growth rates for mass, height, stem area and stem diameter. For the five different traits, we observed four relatively distinct types of response. These responses are summarised in Table \ref{tab:responses} and described further below. The first two traits considered modify behaviour primarily at the start and end of an individuals growth trajectory, and are therefore termed ``ontogenetic traits''. The remaining three traits are termed ``development traits'' because they moderate the rate of movement along an individuals growth trajectory.

@@

\noindent{\bf Seed size, $\omega$:}
Increasing $\omega$ in our model causes seedlings to be larger and fecundity to decrease. As such, the only effect of seed size on growth comes from changing the plant's initial  size. The plots in Fig. \ref{fig:conceptual}, which show changes in growth rate with plant size, also express the expected changes in the growth of seedlings due to changes in seed size. Under similar light conditions, larger seedlings are predicted to have faster absolute growth rates (in diameter, height, and mass), because of their greater total leaf area. At the same time, relative growth rate is predicted to decrease with size, because the ratio of leaf area to support mass decreases with plant size. As plants grow, differences in initial mass will decrease in importance, relative to other factors influencing growth through the life-cycle. We therefore expect the correlations between seed size and growth rate observed for seedlings to disappear among larger plants.

\noindent{\bf Height at maturation, $H_{\rm m}$:}
$H_{\rm m}$ moderates growth by adjusting the amount of energy invested in growth, i.e.~the term $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqns \ref{eq:dhdt} and \ref{eq:dast}. Greater height at maturation ($H_{\rm m}$) can thus lead to a growth advantage by increasing $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ (Fig. \ref{fig:growth_light_height}). At smaller sizes, there is no differentiation among species, because most individuals are focusing on growth. At larger sizes, individuals of some species are allocating a majority of their surplus energy to reproduction, which reduces their growth rate relative to those species with greater $H_{\rm m}$(Fig. \ref{fig:growth_light_height}).

@@

First, our results highlight the importance of allocation decisions and turnover costs in determining growth dynamics. Much of current ecosystem research focusses on  photosynthesis, respiration, and resultant fluxes of carbon, with little attention devoted to allocation and turnover \citep[for comparisons of models see][]{Sitch-2008, DeKauwe-2014}. Yet, for four of the five traits considered here, trait values do not influence photosynthesis, respiration, or net primary production (=photosynthesis - respiration). In fact, the analysis with $\phi$ shows that increased growth rate can occur even at a distinct cost to the plant's carbon budget. Low $\phi$ results in high leaf turnover, such that individuals with a $\phi$ have lower mass production. It is this property that makes them shade intolerant. And yet they can still achieve a growth advantage (when small), because the benefits of cheap leaf construction outweigh the costs of high leaf turnover.

@@

\subsection{A -- Full description of the \texttt{FF16} growth model}\label{sec:ff16}

>>>>

\keywords{growth rate, life-history, ontogeny, size, leaf mass per area, wood density, shade-tolerance, seed mass, height at maturation, leaf nitrogen}

@@

\section{Abstract}

Plant species differ in many functional traits, including aspects of tissue construction and energy allocation, that drive differences in rates of photosynthesis, biomass allocation, and tissue turnover. Yet, it remains unclear how -- and even if -- such traits influence whole-plant growth, with the simple linear relationships predicted by existing theory often lacking empirical support. Here we present a new theoretical framework for understanding the effect of diverse functional traits on plant growth, extending a widely-used theoretical model that links growth rate in seedlings with a single leaf trait to explicitly include influences of size, light environment, and five other prominent traits: seed mass, height at maturation, leaf mass per unit leaf area, leaf nitrogen per unit leaf area, and wood density. Using a model based on energy uptake and allocation, we explain why the influence of prominent traits on growth rate often varies with plant size and why the impact of size varies among traits. Considering a plant's growth rate in height, we find the influence of: i) leaf mass per unit leaf area is strong in small plants but weakens with size, ii) leaf nitrogen per unit leaf area does not change with size, iii) is present across sizes but is strongest at intermediate sizes, iv) height at maturation strengthens with size, and v) seed mass decreases with size. Our theoretical model shows how these changes in trait-growth relationships with size arise via changes in biomass partitioning and reproductive allocation. Moreover, we show how traits moderate plant responses to light environment and also determine shade tolerance, supporting diverse empirical results. By disentangling the effects of plant size, light environment and traits on growth rates, our results provide a solid theoretical foundation for trait ecology and thus provide a platform for understanding growth across diverse species around the world.

@@

One possible reason that theoretical predictions have not been supported is that the effects of traits in existing theory are realised mainly via influences on primary productivity (photosynthesis - respiration) or mass production \citep{Enquist-2007}. By contrast, the physiology of traits such as leaf mass per unit leaf area and wood density suggests they have influence by altering the allocation of energy among different tissues, rather than changing mass production \citep{Falster-2011, Gibert-2016}. A second concern for theory focussed on mass production is that altered mass production is only possible outcome. Measuring mass production is really only practical for small plants that can be easily harvested. For larger plants, growth is more-often measured as increment in either diameter or height. For example, research centres and forestry agencies have been recording stem diameter growth in 1000's of plots across the globe \citep{Purves-2008, Anderson-2015, Kunstler-2016}. Theory considering only primary production is therefore of limited potential use.

In this paper we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth (Table \ref{tab:phenomena}); and thereby offer fresh insight into the way traits influence plant demography throughout across the life-cycle. This model can be viewed as an integration of several related approaches \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011}. Our primary focus in this article is to explain the pattern that has been gradually emerging -- that the effect of traits on plant growth may be modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}. As noted above, the paper by \citet{Gibert-2016} outlined a framework describing how traits influence growth. Here we extend the results of \citet{Gibert-2016} to provide a functioning model and explain, from the point of view of energy uptake and allocation, why it makes sense that the effect of some traits changes with size. We also show how our approach can account for other phenomena, including changes in growth and shade tolerance with traits, individual size, and light environment. Our view is that trait-based approaches -- which focus on differences among species -- should be integrated within a general model of plant growth, also able to capture patterns of growth through ontogeny. Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Herault-2011}. In contrast, basal-area growth continues to increase with size \citep{Sillett-2010, Stephenson-2014}. And all growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}. Shade tolerance also varies among species, correlates with traits \citep{Lusk-2008,Poorter-2006}, and tends to decrease with increasing size \citep{Givnish-1988, Lusk-2008}. These diverse phenomena -- summarised in Table \ref{tab:phenomena} -- are deserving of a comprehensive and integrated explanation.

@@

We assume growth is fundamentally driven by energy income and subsequent distribution throughout the plant. Applying a standard approach, the amount of biomass available for growth, $\frac{{\rm d}B}{{\rm d}t}$ is given by the difference between income (photosynthesis) and losses (respiration and turnover) within the plant \citep{Makela-1997, Thornley-2000} (Fig. \ref{fig:conceptual}a, eqn \ref{eq:dbdt}). Photosynthesis is the product of the average photosynthetic rate per unit leaf area ($\bar{p}$) and total leaf area ($A_{\rm l}$). The constant $y$ in eqn \ref{eq:dbdt} converts mol CO$_2$ into units of biomass. We assume that $\bar{p}$ increases with light but with diminishing returns, and respiration and turnover rates of different tissues are constants that might differ with traits (see below). While many vegetation models use a detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$, e.g. as functions of temperature, such detail is not needed here, as even with detailed models, $\frac{{\rm d}B}{{\rm d}t}$ always comes down to simple subtraction of income and losses.

@@

Eqn \ref{eq:dbdt} can also be used to estimate a measure of shade tolerance, sometimes referred to as the ``whole-plant-light-compensation point'' ({\wplcp}): the light level at which a plant's photosynthetic gains just balance the costs of tissue turnover and respiration \citep{Givnish-1988, Baltzer-2007, Lusk-2013}. In general, average assimilation rate per leaf area $\bar{p}$ increases with light level or canopy openness, $E$. The {\wplcp} can be estimated by solving for the the value of canopy openness $E=E^*$ giving $\frac{{\rm d}B} {{\rm d}t}(E)=0$ (Fig. \ref{fig:wplcp_idea}). From eqn \ref{eq:dbdt}, this occurs when
\begin{equation}\label{eq:wplcp}
\bar{p}(E^*) =\frac{\sum_{i={\rm l,b,s,r}} \, M_{\rm i} \left(\frac{k_{\rm i}}{y} + r_{\rm i}\right)}{A_{\rm l}},
\end{equation}
where $\bar{p}(E)$ is the rate of photosynthesis at canopy openness $E$. The {\wplcp} occurs at the points where the photosynthetic production (per unit leaf area) line intersects with the sum of maintenance and respiration costs (per unit leaf area) for each tissue (Fig. \ref{fig:wplcp_idea}). Traits will influence the {\wplcp} if they effect either income or costs. Also, since the amount of stem support naturally increases with height, the {\wplcp} also naturally increases with height \citep{Givnish-1988} (Fig. \ref{fig:wplcp_idea}).

@@

It worth noting that because eqns \ref{eq:dMtdt}-\ref{eq:dDdt} are derived using the chain rule of differential calculus, they are mathematically true and must therefore hold for any potential growth model. To make explicit predictions via this framework, then requires an explicit model of plant construction and function; i.e. we must put equations behind all there terms in Fig. \ref{fig:conceptual}b.

Here we adopt a model of plant construction and function that can be considered a first-order functional-balance model, similar to those implemented by \citet{Makela-1997} and \citet{Moorcroft-2001}. We could also call it ``isometric'', because the assumptions preserve the proportions among the relevant variables as the plant grows. Table \ref{tab:allometry} provides key equations of the model (see \citet{Falster-2016} for full derivation). In particular, we assume:
\begin{description}
\item[Assumption 1:] A log-log scaling relationship with slope of 0.5 between the leaf area of a plant and its height (eqn \ref{eq:ha}).
\item[Assumptions 2:] A log-log scaling relationship with slope of 1.0 between the leaf area and cross-sectional area of sapwood in a plant (eqn \ref{eq:as}).
\item[Assumptions 3:] A log-log scaling relationship with slope of 1.0 between the leaf area and cross-sectional area of bark and phloem in a plant (eqn \ref{eq:ab}).
\item[Assumptions 4:] A log-log scaling relationship with slope of 1.0 between the leaf area and mass of fine roots in a plant (eqn \ref{eq:mr}).
\item[Assumption 5:] The vertical distribution of leaf within the canopy of an individual plant remains constant through ontogeny.
\end{description}
Assumption 1 accounts for the architectural layout of the plant. Assumptions 2-4 are realisations of the pipe model \citep{Shinozaki-1964}, whereby the cross-sectional area of conducting tissues are proportional to leaf area. To describe the vertical distribution of leaf area within the canopy of an individual plant (assumption 5), we use the model of \citet{Yokozawa-1995}, which can account for a variety of canopy profiles through a single parameter $\eta_c$, varying from 0-1 (for details, see Supplementary Materials).

Combined, the functional-balance assumptions from Table \ref{tab:allometry}a lead directly to equations describing the mass of sapwood and bark in relation to leaf, and the amount of leaf in relation to height (Table \ref{tab:allometry}b). Substituting from Table \ref{tab:allometry} into eqns \ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives all the necessary terms needed to implement the growth model described in Fig. \ref{fig:conceptual}.

An important prediction arising from the eqns in Table \ref{tab:allometry} is that the amount of live stem tissue supporting each unit of leaf area increases linearly with  height:
\begin{equation}\label{eq:msml}
\frac{M_{\rm b} + M_{\rm s}}{A_{\rm l}} = (1+ \alpha_{\rm b1}) \,\theta \, \rho \, \eta_c \, H,
\end{equation}
where $\alpha_{\rm b1}, \theta, \rho, \eta_c$ are all traits (Table \ref{tab:traits}), i.e. properties that are assumed approximately constant through ontogeny.

@@

We now consider how trait-based trade-offs enter into the above growth model. It is essential that any trait include both a benefit and cost in terms of plant function and/or life history; otherwise we would expect never-ending directional trait values towards more beneficial values. For present purposes, we consider five prominent traits for which we can posit specific costs and benefits, outlined in Table \ref{tab:traits}. In postulating potential benefits and costs, we consider only those thought to arise as direct biophysical consequences of varying a trait. The trade-offs are then implemented as follows.
\begin{description}
\item[Seed mass, $\omega$:] There is a direct energetic trade-off between the a plant's fecundity and the size of it's seeds. Moreover, we assume larger seeds result in larger seedlings; the seed-size number tradeoff thereby translates into a demographic tradeoff between the number of seedlings produced by a parent and their initial size.
\item[Height at maturation, $H_{{\rm mat}}$:] This trait moderates the inevitable energetic trade-off between growth and reproduction. We assume the fraction of energy allocated to reproduction, $1- \frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ is determined by $H_{{\rm mat}}$ via a function $r(H, H_{{\rm mat}}) = \frac{r_{\rm r1}} {1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{{\rm mat}}\right)\right)}$, where $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition.
\item[Nitrogen per unit leaf area,  $\nu$:] We allow for the maximum photosynthetic capacity of the leaf to vary with leaf nitrogen per unit area, as $A_{\rm max} = \beta_{{\rm lf1}} \, \left(\frac{\nu}{\nu_0}\right)^{\beta_{{\rm lf5}}}$, where $\beta_{{\rm lf1}}$, $\nu_0$ and $\beta_{{\rm lf5}}$ are constants. Respiration rates per unit leaf area are also assumed to vary linearly with leaf nitrogen per unit area, as $\beta_{{\rm lf4}}\, \nu$.\footnote{Add plot of this relationship, from data in \citet{Wright-2004}.}
\item[Leaf mass per unit area, $\phi$:] This trait directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}). In addition, we link $\phi$ to the rate of leaf turnover ($k_{\rm l}$), based on a widely observed scaling relationship from \citet{Wright-2004}: $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants. The rate of leaf respiration per unit area is assumed independent of $\phi$, as such the mass-based rate is adjusted accordingly whenever $\phi$ is varied.
\item[Wood density, $\rho$:] This trait directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}). In addition, we link $\rho$ to the rate of sapwood and bark turnover, mirroring the relationship assumed for leaf turnover: $k_{\rm b} = k_{\rm s} = \beta_{{\rm ks1}} \, \left(\frac{\rho}{\rho_0}\right)^{-\beta_{{\rm ks2}}}$ where $\beta_{{\rm ks1}}$, $\rho_0$ and $\rho_{{\rm ks2}}$ are empirical constants. The rate of sapwood and bark respiration per unit stem volume is assumed to be independent of $\rho$, as such the mass-based rate is adjusted accordingly  whenever $\rho$ is varied.
 \end{description}

@@

The growth model described above has been implemented as the \texttt{FF16} physiological model within the {\plant} package \citep{Falster-2016} for \textsc{r} \citep{R-2015}. The {\plant} package also makes use of supporting packages \texttt{Rcpp} \citep{Eddelbuettel-2013} and the Boost Library for C++\citep{Schaling-2014}, via the package \texttt{BH} \citep{Eddelbuettel-2015}. To encode the trait-based trade-offs described above, we use the capacity to in {\plant} to provide a ``hyper-parameterisation'' function, which enables various parameters to covary with traits (for full details see Supplementary Materials -- A).

For the most part, parameters used in the current analysis were sourced from \citet{Falster-2016} (see Table \ref{tab:params} for values). The only exceptions are i))parameters effecting the relationships outlined in Table \ref{tab:allometry}a, which were fit using data described below, and ii) parameters describing the function for reproductive allocation. By default, we set $r_{\rm r1}=0.8$ and $r_{\rm r2}=10$, implying a relatively rapid transition to reproduction at $H=H_{{\rm mat}}$ (see panel for $\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ in Fig. \ref{fig:conceptual}).

@@

The functional-balance assumptions outlined in Table \ref{tab:allometry}a are well-supported by the available data. Fig. \ref{fig:assumptions} shows data sourced from the \textsc{baad} \citep{Falster-2015b}, with solid lines indicating {\sma} lines fit to each species in the data. Dashed lines indicate the slope of predictions under functional-balance assumptions in Table \ref{tab:allometry}a. As expected, species differed in elevation, but less so in the slope of the fitted lines, with slopes aligning with those predicted by the functional-balance assumptions.

@@

Our growth model suggests an intrinsically size-dependent pattern of growth, which aligns with well-known empirical patterns (Table \ref{tab:phenomena}). The panels in Fig. \ref{fig:conceptual} show the expected patterns for a typical woody plant, obtained by applying the functional-balance model of function (Table \ref{tab:allometry}). Height growth shows a hump-shaped pattern with size, first increasing then decreasing. This pattern results from systematic changes in the four components of eqn \ref{eq:dhdt} with size, including a strong decline in the fraction of plant that is leaf declines with increasing size (Fig. \ref{fig:conceptual}), and increasing reproductive allocation (Fig. \ref{fig:conceptual}). In contrast, basal-area growth continues to increase with size, due to an increasing influence of stem turnover. Diameter growth shows a weakly hump shaped curve, tapering off slightly at larger sizes, in part because of the allometric conversion from basal area to diameter (eqn \ref{eq:dDdt}, and in part because of increased reproductive allocation in older trees (Fig \ref{fig:conceptual}). All growth measures decrease sharply with size when expressed as relative growth rates (results not shown).

@@

We analysed the response of growth rate to five different traits under the assumed trade-offs (Table \ref{tab:traits}). We considered changes in absolute and relative growth rates for mass, height, stem area and stem diameter. The first two traits considered modify behaviour primarily at the start and end of an individuals growth trajectory, and are therefore termed ``ontogenetic traits''. The remaining three traits are termed ``development traits'' because they moderate the rate of movement along an individuals growth trajectory. Across the five different traits, we observed four relatively distinct types of response. These responses are summarised in Table \ref{tab:responses} and described further below.

@@

\noindent{\bf Seed size, $\omega$.}
Increasing $\omega$ in our model causes seedlings to be larger and fecundity to decrease. As such, the only effect of seed size on growth comes from changing the plant's initial  size. The plots in Fig. \ref{fig:conceptual}, which show changes in growth rate with plant size, also express the expected changes in the growth of seedlings due to changes in seed size. Under similar light conditions, larger seedlings are predicted to have faster absolute growth rates in all metrics  because of their greater total leaf area. At the same time, relative growth rate is predicted to decrease with size, because the ratio of leaf area to support mass decreases with plant size. As plants grow, differences in initial mass will decrease in importance, relative to other factors influencing growth through the life-cycle. We therefore expect the correlations between seed size and growth rate observed for seedlings to disappear among larger plants.

\noindent{\bf Height at maturation, $H_{{\rm mat}}$.}
$H_{{\rm mat}}$ moderates growth by adjusting the amount of energy invested in growth, i.e.~the term $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqns \ref{eq:dhdt} and \ref{eq:dast}. Greater $H_{{\rm mat}}$ can thus lead to a growth advantage by increasing $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ (Fig. \ref{fig:growth_light_height}). At smaller sizes, there is no differentiation among species, because most individuals are focusing on growth. At larger sizes, individuals of some species are allocating a majority of their surplus energy to reproduction, which reduces their growth rate relative to those species with greater $H_{{\rm mat}}$ (Fig. \ref{fig:growth_light_height}).

@@

First, our results highlight the importance of allocation decisions and turnover costs in determining growth dynamics. Much of current ecosystem research focusses on factors effecting primary production -- photosynthesis, respiration, and resultant fluxes of carbon -- with little attention devoted to allocation and turnover \citep[for comparisons of models see][]{Sitch-2008, DeKauwe-2014}. Yet, for four of the five traits considered here, trait values do not influence net primary production. In fact, the analysis with $\phi$ shows that increased growth rate can occur even at a distinct cost to the plant's carbon budget. Low $\phi$ results in high leaf turnover, such that individuals with a $\phi$ have lower mass production. It is this property that makes them shade intolerant. And yet they can still achieve a growth advantage (when small), because the benefits of cheap leaf construction outweigh the costs of high leaf turnover.

@@

\noindent A -- Additional details on the \texttt{FF16} growth model

\noindent B -- Supplementary Tables

\noindent C -- Supplementary Figures
